---
title: "Weather and Corn Yield Regressions"
author: "Anna Marshall"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(R.matlab)
library(rnassqs)
```

## Overview
The USDA National Agricultural Statistics Service (NASS) provides a data repository of U.S. agricultural production. Learn more about it here. (https://www.nass.usda.gov/) Here we look at corn and soybean yields at different timescales, locations, and temperatures. 

## Methods
PRISM daily maximum temperature data and NASS crop yield data was downloaded and brought into R for analyses. The methods for analysis here focus on multiple regressions. 
```{r tmax data, include=FALSE}
##Weather Data Analysis
### Load the PRISM daily maximum temperatures
# daily max temperature
# dimensions: counties x days x years
prism <- readMat("data/prismiowa.mat")

# look at county #1
t_1981_c1 <- prism$tmaxdaily.iowa[1,,1]
t_1981_c1[366]
plot(1:366, t_1981_c1, type = "l")

ggplot() +
  geom_line(mapping = aes(x=1:366, y = t_1981_c1)) +
  theme_bw() +
  xlab("day of year") +
  ylab("daily maximum temperature (°C)") +
  ggtitle("Daily Maximum Temperature, Iowa County #1")


```
```{r, include=FALSE}
#Tidying up
# assign dimension names to tmax matrix
dimnames(prism$tmaxdaily.iowa) <- list(prism$COUNTYFP, 1:366, prism$years)

# converted 3d matrix into a data frame
tmaxdf <- as.data.frame.table(prism$tmaxdaily.iowa)

# relabel the columns
colnames(tmaxdf) <- c("countyfp","doy","year","tmax")
tmaxdf <- tibble(tmaxdf)

```



```{r, include=FALSE}
## Temperature trends

### Summer temperature trends: Winneshiek County
tmaxdf$doy <- as.numeric(tmaxdf$doy)
tmaxdf$year <- as.numeric(as.character(tmaxdf$year))

winnesummer <- tmaxdf %>%
  filter(countyfp==191 & doy >= 152 & doy <= 243) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnesummer, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_summertmax <- lm(meantmax ~ year, winnesummer)
summary(lm_summertmax)

```

```{r, include=FALSE}
### Winter Temperatures - Winneshiek County
winnewinter <- tmaxdf %>%
  filter(countyfp==191 & (doy <= 59 | doy >= 335) & !is.na(tmax)) %>%
  group_by(year) %>%
  summarize(meantmax = mean(tmax))

ggplot(winnewinter, mapping = aes(x = year, y = meantmax)) +
  geom_point() +
  theme_bw() +
  labs(x = "year", y = "Tmax (°C)") +
  geom_smooth(method = lm)

lm_wintertmax <- lm(meantmax ~ year, winnewinter)
summary(lm_wintertmax)

```


```{r, include=FALSE}
### Multiple regression -- Quadratic time trend

winnewinter$yearsq <- winnewinter$year^2

lm_wintertmaxquad <- lm(meantmax ~ year + yearsq, winnewinter)
summary(lm_wintertmaxquad)
winnewinter$fitted <- lm_wintertmaxquad$fitted.values

ggplot(winnewinter) +
  geom_point(mapping = aes(x = year, y = meantmax)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "tmax")

```


```{r, include=FALSE}
### Download NASS corn yield data

# set our API key with NASS
nassqs_auth(key = "9DAB6678-0A6B-3645-8E43-F2DCBB5F659C")

# parameters to query on 
params <- list(commodity_desc = "CORN", util_practice_desc = "GRAIN", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
cornyieldsall <- nassqs_yields(params)

cornyieldsall$county_ansi <- as.numeric(cornyieldsall$county_ansi)
cornyieldsall$yield <- as.numeric(cornyieldsall$Value)

# clean and filter this dataset
cornyields <- select(cornyieldsall, county_ansi, county_name, yield, year) %>%
  filter(!is.na(county_ansi) & !is.na(yield))
cornyields <- tibble(cornyields)

```

## Results and Discussion

### Question 1a: Extract Winneshiek County corn yields, fit a linear time trend, make a plot. Is there a significant time trend?

Based on the fit of the linear time trend we can see that there is a significant trend between corn yield and time (p-value less than our alpha of 0.05). As years have increased, corn yield has also increased. We can also see this in the high R2 value of 0.75, suggesting a good linear model fit. 

```{r, message=FALSE, results=FALSE}
winnecorn <- cornyields %>%
  filter(county_name== 'WINNESHIEK') %>%
  group_by(year) %>%
  summarize(meanyield = mean(yield))
ggplot(winnecorn, mapping = aes(x = year, y = meanyield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Year", y = "Corn Yield (Bushels per Acre)") +
  geom_smooth(method = lm)

lm_winnecorn <- lm(meanyield ~ year, winnecorn)
summary(lm_winnecorn)
```


### Question 1b: Fit a quadratic time trend (i.e., year + year^2) and make a plot. Is there evidence for slowing yield growth? 
The quadratic time trend fits the dataset well as evident by a R2 value of 0.75. However, the trend is positive and appears to suggest that the yield growth is still increasing. 
```{r, message=FALSE}
winnecorn$yearsq <- winnecorn$year^2

lm_winnequad <- lm(meanyield ~ year + yearsq, winnecorn)
summary(lm_winnequad)
winnecorn$fitted <- lm_winnequad$fitted.values

ggplot(winnecorn) +
  geom_point(mapping = aes(x = year, y = meanyield)) +
  geom_line(mapping = aes(x = year, y = fitted)) +
  theme_bw() +
  labs(x = "year", y = "Yield (Bushels per acre)")
```

### Question 2 -- Time Series: Let's analyze the relationship between temperature and yields for the Winneshiek County time series. Use data on yield and summer avg Tmax. Is adding year or Tmax^2 to your model helpful? Make a plot and interpret the results.

Here, adding Tmax^2 to our model does not help our fit. We have a R2 of 0.2, which suggests a poor model fit and lack of a relationship between temperature and yields for Winneshiek County. 

```{r, message=FALSE}
#bring in sumemr temp into yield
winnetime<-inner_join(winnecorn,winnesummer)
#now quadratic with temp data
winnetime$tempsq<- winnetime$meantmax^2
lm_winnetime_quad<- lm(meanyield~meantmax +tempsq,winnetime)
summary(lm_winnetime_quad)
winnetime$quadfitted<- lm_winnetime_quad$fitted.values
#plotting regression
ggplot(winnetime)+
  geom_point(mapping=aes(x=meantmax,y=meanyield))+
  geom_line(mapping=aes(x=meantmax,y=quadfitted))+
  labs(x='Temperature (C)',y='Yield(Bushels per acre)')
```


### Question 3 -- Cross-Section: Analyze the relationship between temperature and yield across all counties in 2018. Is there a relationship? Interpret the results.
Here, we do not see a strong linear relationship between temperature and yield across all counties in 2018. This is evident by the lack of a clear visual linear trend in the plot below. 
```{r}
#gives us yield for all counties in 2018
yield2018<- cornyieldsall %>%
  filter(year==2018) %>%
  group_by(county_name) %>%
  unique() %>%
  filter(!is.na(county_ansi))
#gives max summer temp for 2018 per county
county_summer<- tmaxdf %>%
  group_by(countyfp) %>%
  filter(year==2018) %>%
  filter(doy>=152 & doy<=243) %>%
  summarize(meantmax=mean(tmax)) %>%
  rename(county_ansi="countyfp")
#change from factor to numeric
county_summer$county_ansi<- as.numeric(as.character(county_summer$county_ansi))
#join data frames together
county_summer_yield<- left_join(yield2018,county_summer,by="county_ansi")

#plot data
ggplot(county_summer_yield, aes(x=yield, y=meantmax))+
  geom_point()+
  geom_smooth(mapping = aes(x = yield, y = meantmax),method = lm) +
  theme_bw()+
  labs(x="Yield (Bushels per acre)", y="Temperature (C)")
```

### Question 4 -- Panel: One way to leverage multiple time series is to group all data into what is called a "panel" regression. Convert the county ID code ("countyfp" or "county_ansi") into factor using as.factor, then include this variable in a regression using all counties' yield and summer temperature data. How does the significance of your temperature coefficients (Tmax, Tmax^2) change? Make a plot comparing actual and fitted yields and interpret the results of your model.

Here, once again a R2 value of 0.65 suggests the model predicts the actual yield well. The data points are more closely correlated between fitted and actual yields with higher fitted and actual yield values. 
```{r, include=FALSE}
yieldall<- cornyieldsall %>%
  group_by(county_name) %>%
  unique() %>%
  filter(!is.na(county_ansi))
county_summer_yield2<- left_join(yieldall,county_summer,by="county_ansi")

#bringing in data from Q3 and making county_ansi a factor
county_summer_yield2$county_ansi<- as.factor(county_summer_yield2$county_ansi)
head(county_summer_yield2)
#squaring the tmax data
county_summer_yield2$meantmaxsq<- county_summer_yield2$meantmax^2
lm_county_summer_yield2<- lm(yield~meantmax+meantmaxsq+year+county_ansi,county_summer_yield2)
summary(lm_county_summer_yield2)
county_summer_yield2$fitted <- lm_county_summer_yield2$fitted.values
#now plotting the data
ggplot(county_summer_yield2, aes(x=yield, y=meantmax))+
  geom_point()+
  #geom_line(mapping=aes(x=yield,y=fitted))+
  theme_bw()+
  labs(x="Yield (Bushels per acre)", y="Temperature (C)")

```

```{r, message=FALSE, results=FALSE}
# Want yield for all counties, all years:
county_yield_panel <- cornyieldsall %>% 
  group_by(county_name) %>% 
  unique() %>% 
  filter(!is.na(county_ansi))
#Want maximum summer temperatures per county. 
county_summer_panel <- tmaxdf %>%
  group_by(countyfp) %>%
  filter(doy >= 152 & doy <= 243) %>%
  summarize(meantmax = mean(tmax)) %>% 
  rename(county_ansi = "countyfp")
county_summer_panel$county_ansi <- as.numeric(as.character(county_summer_panel$county_ansi))
#Joining the two dfs together
county_summer_yield_panel <- left_join(county_yield_panel,county_summer_panel, by='county_ansi') 
# combined yield and summer temperature df for all counties
county_summer_yield_panel <- subset(county_summer_yield_panel, select = c(county_ansi, yield, meantmax, year))
#making the county_ansi into a factor:
county_summer_yield_panel$county_ansi <- as.factor(county_summer_yield_panel$county_ansi)
#confirming it is a factor
str(county_summer_yield_panel$county_ansi)
# Squares the tmax value
county_summer_yield_panel$meantmaxsq <- county_summer_yield_panel$meantmax^2
lm_county_summer_yield_panel <- lm(yield ~ county_ansi + meantmax + meantmaxsq + year, county_summer_yield_panel)
summary(lm_county_summer_yield_panel)
#str(county_summer_yield)
#head(county_summer_yield_panel)
county_summer_yield_panel$fitted <- lm_county_summer_yield_panel$fitted.values
#plotting the df with the linear and fitted model
ggplot(county_summer_yield_panel) +
  geom_point(mapping = aes(x = yield, y = fitted)) +
  geom_smooth(mapping = aes(x = yield, y = fitted),method = lm) +
  labs(x = "Actual Yield (Bushels per Acre)", y = "Fitted Yield")
```

### Question 5 -- Soybeans: Download NASS data on soybean yields and explore either a time series relationship for a given county, the cross-sectional relationship for a given year, or a panel across all counties and years.

Here, I have plotted the soybean yield on an annual basis. There appears to be a linear relationship with an increase in soybean yield each year, as evident by an R2 value of 0.65.  
```{r, message=FALSE, results=FALSE}
# set our API key with NASS
nassqs_auth(key = "9DAB6678-0A6B-3645-8E43-F2DCBB5F659C")
?rnassqs
# parameters to query on 
params2 <- list(commodity_desc = "SOYBEANS", statisticcat_desc="YIELD", prodn_practice_desc = "ALL PRODUCTION PRACTICES", year__GE = 1981, state_alpha = "IA")

# download
soyyieldsall <- nassqs_yields(params2)

soyyieldsall$county_ansi <- as.numeric(soyyieldsall$county_ansi)
soyyieldsall$Value <- as.numeric(soyyieldsall$Value)

# clean and filter this dataset
soyyields <- select(soyyieldsall, county_ansi, county_name, Value, year) %>%
  filter(!is.na(county_ansi) & !is.na(Value))
soyyields <- tibble(soyyields)

#now plotting for linear trend line
soylm <- soyyields %>%
  group_by(year) %>%
  summarize(meanyield = mean(Value))
ggplot(soylm, mapping = aes(x = year, y = meanyield)) +
  geom_point() +
  theme_bw() +
  labs(x = "Year", y = "Soybean Yield (Bushels per Acre)") +
  geom_smooth(method = lm)

lm_soylm <- lm(meanyield ~ year, soylm)
summary(lm_soylm)
```


